\papertitle

### 2012-10-12 manuscript.tex
Filled in section on BeagleBoard, minor edits

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper MAV}
% humphrey
In order to conduct these experiments on cooperation we designed a new
flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
RC flyer power train and clap-fling wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
servo-controlled tail elevator provide control in the yaw and pitch axes. The 
on-board ImageProc 2.4 electronics package consists of a 40 MIPS microprocessor, 
6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by 
a 90 mAh lithium polymer battery\footnote{Expand! Accredit it to Stan Baek!}. In routine flight, the H$^2$Bird 
averages 1.2 m/s ground speed and can operate for approximately 10 minutes.
%TODO footnote for Silverlit I-Bird
%TODO footnote for ImageProc

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
%TODO ryan

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO ryan

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/process_flow.pdf}
\caption{Overview of the cooperative control system.}
\label{fig:process_flow}
\end{figure}

%----------------------------------------------------------------------------%
\subsection{Ornithoper Flight Control}

\begin{figure}[tb]
\centering
\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
\caption{H$^2$Bird Ornithopter with flight control axes}
\label{fig:h2Bird_axes}
\end{figure}
%TODO check ibird labels

Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
acceptable tracking performance. We estimate the vehicle pose by naively
integrating the IMU, and use PID for controlling the vehicle.
%TODO Cite Stan paper
One behavior of the H$^2$Bird that separates it from other MAVs is the
high pitch angles achieved in routine flight. Baek\cite{baek:tracking} 
demonstrated control on a similar vehicle using PID and an Euler angle
parameterization of orientation. However, Euler angles suffer from singularities
around high pitch angles, so we instead represent the vehicle orientation
with quaternions. An additional benefit of the quaternion representation is 
the ability to represent relative body and world coordinate angular displacements 
compactly with quaternion multiplication

Our implementation of quaternion-based control is similar to Knoebe's\cite{knoebe:quatcontrol}: 
Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
required to reach the reference pose from the measured pose. We then 
convert this error quaternion to an angle axis representation which is given
as an input to the PID controller.

\begin{equation}
\label{quat_error}
q_e = q'\otimes q_r
\end{equation}
\begin{equation}
\label{quat_angle}
\alpha = 2acos(q_{e,w})
\end{equation}
\begin{equation}
\label{quat_linearize}
Q_e = \alpha*q_e/sin(\alpha /2)
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
%TODO ryan
The task we chose to test our cooperative visual servoing system with is a 
simple target seeking one: the H$^2$Bird must fly through a specified obstacle,
in this case, a wooden frame meant to simulate a window. As the H$^2$Bird does
cannot detect the target on its own, it depends on remote guidance from the
base station. Ground truth data was collected via a motion capture system in
the test room. A sketch of the experimental setup is shown below in Figure 
~\ref{fig:experiment_cartoon}.

% TODO clarify or roll into paragraph
Trials were run as follows:
\begin{enumerate}
\item The window is identified by a user. 
\item The H$^2$Bird is released in view of the camera by hand.
\item The H$^2$Bird attempts to fly to the window with remote guidance from
the base station.
\end{enumerate}

% TODO Fill in test # per battery
Variations in the H$^2$Bird hand launch cause the system to respond differently 
for otherwise identical initial initial conditions. In addition, the H$^2$Bird 
battery was changed approximately every [10] tests, adding more variation to
the H$^2$Bird performance by affecting the motor power output. The effects of 
these noisy factors is considered in Section ~\ref{sec:performance}.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
\caption{Conceptual sketch of experimental environment.}
\label{fig:experiment_cartoon}
\end{figure}
%TODO label experiment cartoon

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}
\label{sec:performance}
%TODO cameron
To bound the performance of our system described in the previous sections, we
determined the feasible set of initial conditions that resulted in successful
window traversal. We then conducted experiments to calculate the success rate
both within the feasible region and on the edges of the feasible region. The
following sections detail how we calculated the feasible and infeasible
regions using system identification and present the results of our
experimental trials.

%----------------------------------------------------------------------------%
\subsection{Ornithopter Flight Control}
\label{sec:flight_control}
%TODO cameron
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
\caption{H$^2$Bird step response}
\label{fig:step_response}
\end{figure}
%TODO reduce number of sig figs

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. We depict the results of this experiment in
Figure~\ref{fig:step_response}. We gathered the data for the experiment by 
integrating the output of H$^2$Bird's on-board gyroscope data. We used MATLAB to
fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
turning behavior to the step response with the physical heading as the state
and the heading set point as the input. This model serves as a simplification
of both the turning dynamics of the ornithopter and the on-board PID
controller regulating the actuator inputs to reach the desired orientation. We
use this model in Section~\ref{sec:visual_servoing} to determine the backwards
reachable region for navigating through the window in simulation.

The step response rise time is approximately 1.4 seconds. The H$^2$Bird
flies at an average forward velocity of 1.2 m/s, so the estimated minimum
turning radius of the robot is 1.07 meters. In experiments, we choose a flight
speed lower than the maximum speed of the ornithopter, to facilitate more
robust tracking.
%TODO footnote for MATLAB

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
\label{sec:visual_servoing}
%TODO cameron

%TODO make these span the page, with identical scales/style
\begin{figure*}[tb]

>>>>

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper MAV}
% humphrey
In order to conduct these experiments on cooperation, we designed a new
flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan
is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
servo-controlled tail elevator provide control in the yaw and pitch axes. The 
on-board electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 
a ground speed of 1.2 m/s and operates for approximately 10 minutes.
%TODO footnote for Silverlit I-Bird
%TODO footnote for ImageProc

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
% ryan
Cooperative behavior is a promising strategy for autonomous navigation of 
unexplored and unstructured environments. We intend that this system will be 
fully mobile in these environments, not tied to a static ground station or 
motion capture laboratory. To demonstrate the feasibility of this approach, 
we designed our ground station to conform to the power and weight 
constraints of highly mobile millirobotic ground vehicles, such as 
OctoRoACH, which can traverse rough terrains with speed and agility [CITATION]. 
%TODO OctoRoACH citation

Rather than power-intensive PC processors, we used the ARM-based 
BeagleBoard single-board computer as reference computation platform. The
BeagleBoard consumes approximately 1 watt while running our computer vision 
and control algorithms. It uses the same processor as popular sub-10 gram 
processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird using a 
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

Our ground station software is based on the Ubuntu Linux distribution, 
modified with a custom version of the Linux kernel. We use the Python 
programming language for control, robot communication, and telemetry, and 
the OpenCV computer vision libraries for image capture and processing.
%TODO footnotes: Ubuntu, Linux, Python, OpenCV
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Concept and Algorithms}
%TODO ryan

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/process_flow.pdf}
\caption{Overview of the cooperative control system.}
\label{fig:process_flow}
\end{figure}
%TODO h2bird label in diagram should have superscript

%----------------------------------------------------------------------------%
\subsection{Ornithoper Attitude Control}

\begin{figure}[!tb]
\centering
\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
\caption{H$^2$Bird ornithopter with attitude control axes and labeled control surfaces.}
\label{fig:h2Bird_axes}
\end{figure}
%TODO label h2bird drawing

Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
maintain high-rate attitude control. We estimate the vehicle pose by naively
integrating the IMU, and use PID for controlling the vehicle.

The high pitch angles H$^2$Bird achieves in routine flight distinguishes it 
from other MAVs. Euler angle parametrizations of the vehicle orientation 
suffer from singularities around these high pitch angles, so we instead 
represent the vehicle orientation with quaternions. An additional benefit of 
the quaternion representation its ability to represent relative body and 
world coordinate angular displacements compactly with quaternion 
multiplication.

Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
required to reach the reference pose from the measured pose (Eq.~\ref{eq:quat_error}). 
\begin{equation}
\label{eq:quat_error}
q_e = q'\otimes q_r
\end{equation}
We then convert this error quaternion to an angle axis, and use this 
representation as input to the PID controller 
(Eqs.~\ref{eq:quat_angle} and~\ref{eq:quat_linearize}). 
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}
\begin{equation}
\label{eq:quat_linearize}
Q_e = \alpha*q_e/\sin(\alpha /2)
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
%TODO humphrey

\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
\caption{Conceptual sketch of experimental environment.}
\label{fig:experiment_cartoon}
\end{figure}
%TODO label experiment cartoon

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}
%TODO cameron
To bound the performance of the cooperative control system, we determine 
the feasible set of initial conditions that result in successful narrow 
passage traversal using both analytical and empirical methods. We conducted 
experiments to validate this model.

The following sections detail how we calculated the feasible and infeasible
regions using system identification and present the results of our
experimental trials.

%----------------------------------------------------------------------------%
\subsection{Ornithopter Attitude Control}
\label{sec:flight_control}
%TODO cameron
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
\caption{H$^2$Bird step response}
\label{fig:step_response}
\end{figure}
%TODO reduce number of sig figs

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. We depict the results of this experiment in
Figure~\ref{fig:step_response}.

We gathered the data for the experiment by integrating the output of 
H$^2$Bird's on-board gyroscope data. We then used MATLAB [FOOTNOTE] to fit a 
simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the turning 
behavior to the step response with the physical heading as the state and the 
heading set point as the input. This model serves as a simplification of 
both the turning dynamics of the ornithopter and the on-board PID-based 
attitude controller. We use this model in Section~\ref{sec:visual_servoing} 
to determine the backwards reachable region for navigating through narrow passages.
%TODO H(s) in a labeled equation
%TODO footnote for MATLAB

The step response rise time is approximately 1.4 seconds. The H$^2$Bird
flies at an average forward velocity of 1.2 m/s, so the estimated minimum
turning radius of the robot is 1.07 meters. In experiments, we choose a flight
speed lower than the maximum speed of the ornithopter, to facilitate more
robust tracking.
%TODO equation(s) for how you arrive at the minimum turning radius
%TODO footnote for MATLAB

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
%TODO cameron

\subsubsection{Reachability Model}
\label{sec:visual_servoing}
%TODO make these span the page, with identical scales/style
\begin{figure*}[tb!]
